---
title: "p8105_hw3_jh4617"
author: "Junjie Hu"
date: "2023-10-12"
output: github_document
editor_options: 
  chunk_output_type: inline
---

```{r load data }
library(p8105.datasets)
library(tidyverse)
data("instacart")

```

# Problem 1

```{r echo = FALSE}
instacart

names(instacart)
```

There are `r length(unique(pull(instacart, aisle)))` aisles. The `r mode(pull(instacart, aisle))` aisle is the most items ordered from.

```{r plot item}
instacart |>
  group_by(aisle) |>
  summarize(sum_order_number = sum(order_number)) |>
  filter(sum_order_number > 10000) |>
  ggplot(aes(x = aisle, y = sum_order_number)) +
  geom_boxplot() +
    labs(
    title = "The number of items ordered in each aisle",
    x = "Aisle",
    y = "Number") 
```

```{r}
instacart |>
  group_by(aisle, product_name) |>
  summarize(n_obs = n()) |>
  arrange(aisle, desc(n_obs)) |>
  top_n(3, n_obs)
```

```{r}
instacart |>
  group_by(product_name, order_dow) |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  summarize(mean_order_hour_of_day = mean(order_hour_of_day)) |>
  mutate(order_dow = recode(order_dow, "0" = "sunday",
                            "1" = "monday",
                            "2" = "tuesday",
                            "3" = "wednesday",
                            "4" = "thursday",
                            "5" = "friday",
                            "6" = "saturday"))|>
  pivot_wider(
    names_from = order_dow,
    values_from = mean_order_hour_of_day
  ) |>
  knitr:: kable()
  
```

# Problem 2

```{r echo = FALSE}
library(p8105.datasets)
data("brfss_smart2010")

```

data cleaning
```{r}
problem2 = 
  brfss_smart2010 |> 
  janitor::clean_names() |>
  filter(topic == "Overall Health",
         response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) |>
    mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))
  
```


2002
```{r}
problem2_2002 =
  problem2 |>
  group_by(locationabbr) |>
  filter(year == "2002") |>
  summarize(n_obs = n_distinct(locationdesc)) |>
  filter(n_obs >= 7)
print(problem2_2002)
```
In 2002, there are `r nrow(problem2_2002)` states were observed at 7 or more locations, which are `r problem2_2002$locationabbr`.

2010
```{r}
problem2_2010 =
  problem2 |>
  group_by(locationabbr) |>
  filter(year == "2010") |>
  summarize(n_obs = n_distinct(locationdesc)) |>
  filter(n_obs >= 7)
print(problem2_2010)
```
In 2010, there are `r nrow(problem2_2010)` states were observed at 7 or more locations, which are `r problem2_2010$locationabbr`.


```{r}
problem2_ggplot = 
  problem2 |>
  filter(response == "Excellent") |> 
  group_by(locationabbr, year) |>
  summarize(mean_data_value = mean(data_value, na.rm = TRUE)) 

ggplot(data = problem2_ggplot, aes( x = year, y = mean_data_value, group = locationabbr, color = locationabbr)) +
  geom_point() +
  geom_line() +
   labs(x = "Year", 
        y = "Mean Data Value", 
        title = "Mean Data Value for Excellent Response by Year") +
  theme_minimal()
```

The mean data value of states is fluctuating across time. The mean data value across state is quite different too. The highest mean data value is `r max(pull(problem2_ggplot, mean_data_value))` in `r problem2_ggplot |> filter(mean_data_value == "29.46") |> pull(locationabbr)` when `r problem2_ggplot |> filter(mean_data_value == "29.46") |> pull(year)`. The lowest mean data value is `r min(pull(problem2_ggplot, mean_data_value))` in `r problem2_ggplot |> filter(mean_data_value == "11.5") |> pull(locationabbr)` when `r problem2_ggplot |> filter(mean_data_value == "11.5") |> pull(year)`. 

```{r}
problem2_ggplot2 = 
  problem2 |>
  group_by(locationabbr, locationdesc, response) |>
  filter(locationabbr == "NY" & year %in% c(2006, 2010)) 
  
ggplot(data = problem2_ggplot2, aes(x = response, y = data_value)) +
  geom_boxplot() +
  facet_grid(~year) + 
  labs( y = "Data Value", 
        title = "The distribution of data value for responses in NY state")

ggplot(data = problem2_ggplot2, aes(x = data_value, fill = response)) +
  geom_density(alpha = .5) +
  facet_grid(~year) + 
  labs( x = "Data Value", 
        y = "probability",
        title = "The distribution of data value for responses in NY state")
```

```{r echo = FALSE}
problem2_poor_2006 = problem2_ggplot2 |> filter(response == "Poor" & year == "2006") 

problem2_poor_2010 = problem2_ggplot2 |> filter(response == "Poor" & year == "2010") 

problem2_fair_2006 = problem2_ggplot2 |> filter(response == "Fair" & year == "2006") 

problem2_fair_2010 = problem2_ggplot2 |> filter(response == "Fair" & year == "2010") 

problem2_good_2006 = problem2_ggplot2 |> filter(response == "Good" & year == "2006") 

problem2_good_2010 = problem2_ggplot2 |> filter(response == "Good" & year == "2010") 

problem2_verygood_2006 = problem2_ggplot2 |> filter(response == "Very good" & year == "2006")

problem2_verygood_2010 = problem2_ggplot2 |> filter(response == "Very good" & year == "2010")

problem2_excellent_2006 = problem2_ggplot2 |> filter(response == "Excellent" & year == "2006") 

problem2_excellent_2010 = problem2_ggplot2 |> filter(response == "Excellent" & year == "2010") 
```

The median data value of people who responded "poor" is the lowest among all response both in 2006 and 2010, which is equal to `r median(pull(problem2_poor_2006, data_value))` in 2006, and `r median(pull(problem2_poor_2010, data_value))` in 2010. The median data value of people who responded "fair" is the second lowest among all response, which is equal to `r median(pull(problem2_fair_2006, data_value))` in 2006, and `r median(pull(problem2_fair_2010, data_value))` in 2010. The median data value of people of people responded "Good" is equal to `r median(pull(problem2_good_2006, data_value))` in 2006, and `r median(pull(problem2_good_2010, data_value))`. The median data value of people of people responded "Very Good" is equal to `r median(pull(problem2_verygood_2006, data_value))` in 2006, and `r median(pull(problem2_verygood_2010, data_value))`. The median data value of people of people responded "Excellent" is equal to `r median(pull(problem2_excellent_2006, data_value))` in 2006, and `r median(pull(problem2_excellent_2010, data_value))`.


# Problem 3

```{r}
library(tidyverse)

covar = 
  read.csv("nhanes_covar.csv", skip = 2) |> 
  janitor::clean_names() |> 
  magrittr::set_colnames(c("seqn","sex","age","bmi","education")) |>
  slice(-2) 

accel = 
  read.csv("nhanes_accel.csv") |> 
  janitor::clean_names() |> 
  mutate(seqn = as.character(seqn))


problem3_join = 
  left_join(covar, accel) |> 
  filter(age >= 21) |> 
  drop_na(seqn | sex | age | bmi | education) |> 
  mutate(sex = recode(sex, "1" = "Male", "2" = "Female")) |>
  mutate(education = factor(education, levels = c( "1", "2", "3"))) |>
  mutate(education = recode(education, "1" = "Less than high school", "2" = "High school equivalent", "3" = "More than high school")) 
   

```

```{r}
problem3_join |>
  group_by(sex, education) |>
  summarize(n_obs = n()) |>
  pivot_wider(
    names_from = education,
    values_from = n_obs ) |>
knitr:: kable()

problem3_join |>
  mutate(age = as.numeric(age)) |>
  arrange(sex, education) |>
  ggplot(aes(x = sex, y = age)) +
  geom_boxplot() +
  facet_grid(.~education) +
  labs( x = "Sex", 
        y = "Age",
        title = "The age distribution for men and women in each education category")
```

In `Less than high school` category, the age of female and male are similar. In `High school equivalent` category, the age of female is higher than male's. In `More than high school` category, the age of female is slightly lower than male.

```{r}
total_activity = 
  problem3_join |>
  mutate(age = as.numeric(age)) |>
  pivot_longer(
    starts_with("min"),
    names_to = "minute",
    values_to = "value") |>
  group_by(seqn, age, sex, education) |>
  summarize(sum_min = sum(value))

ggplot(data = total_activity, aes(x = age, y = sum_min, color = sex)) +
  geom_point(alpha = 0.3) +
  geom_smooth(aes(group = sex), se = FALSE) +
  facet_grid(.~education) + 
  labs(
    x = "Age",
    y = "Total_Activity") 
```

In `Less than high school` category, female have a higher total acitivity than male in the early age, however, in the older age, the total activity of male is higher than female. In `High school equivalent` category, except at the very early of the age, the total activity of female is generally higher than male. In `More than high school` category, the total activity of female is always higher than male.

```{r}
problem3_plot =
  problem3_join |>
  pivot_longer(
    starts_with("min"),
    names_to = "minute",
    values_to = "value" ) |>
  mutate(minute = sub("min", "", minute))|>
  mutate(minute = as.numeric(minute))

ggplot(data = problem3_plot, aes(x = minute  , y = value, color = sex)) +
  geom_smooth(aes(group = sex)) +
  facet_grid(.~education) +
  labs(
    title = "The activity value among time between male and female by education",
    x = "acticity time",
    y = "activity value"
  )


```

In `less than high school` category, the activity value of male and female are generally similar among time. In `high school equivalent` category, the activity value of male and female is similar before 500 minute and after 1000 minute, however female's activity value is slightly higher than male from 500 to 1000 minute. In `more than high school` category, the activity value of male and female is similar before 500 minute and after 1000 minute, however female's activity value is higher than male from 500 to 1000 minute.




